### 2023-11-20 chapters/paper_retrospective.tex
Update retrospective paper according to final submission version

\section*{Abstract}
Effective stroke treatment depends on recognition by call-takers at prehospital telehealth services. This study aimed to develop and assess the potential of machine learning in improving prehospital stroke recognition during medical helpline calls. We used calls from 1 January 2015 to 31 December 2020 in Copenhagen to develop a machine learning-based classification pipeline. Calls from 2021 were used for testing. Calls were transcribed using an automatic speech recognition model and then categorized as stroke or non-stroke using a text classification model. Call-takers achieved a sensitivity of 52.7\% (95\% confidence interval 49.2-56.4\%) with a positive predictive value (PPV) of 17.1\% (15.5-18.6\%). The machine learning framework performed significantly better (p < 0.0001) with a sensitivity of 63.0\% (62.0-64.1\%) and a PPV of 24.9\% (24.3-25.5\%). Thus, a machine learning framework for recognising stroke in prehospital medical helpline calls may become a supportive tool for call-takers, aiding in early and accurate stroke recognition.

\section{Introduction}
Stroke is a leading cause of disability and death worldwide \parencite{cite1,cite2,cite3}. Effective treatment is time-sensitive, and an optimal outcome is more likely when treatment is administered within the first four and a half hours from stroke onset \parencite{cite4,cite5}. The gateway to ambulance transport and hospital admittance is through prehospital telehealth services, including emergency medical call centres, nurse advice call lines, and out-of-hours health services. In the pre-hospital setting, the use of mobile stroke units has made it possible to deliver advanced treatment faster. As the mobile stroke unit is only dispatched to patients with a suspected stroke, the impact of mobile stroke units is directly influenced by accurate call-taker recognition of stroke \parencite{cite6,cite7}. Call-takers who can rapidly and accurately recognize stroke are therefore crucial in facilitating prompt care in both pre-hospital and in-hospital settings.

@@

We used call recordings and registry data from the Copenhagen Emergency Medical Services (CEMS) and the Danish Stroke Registry (DanStroke)\parencite{cite16} from 2015 to 2020. The call recordings were obtained from two call lines: the 1-1-2 emergency line and the medical helpline 1813 (MH-1813). These data were used to build a machine learning framework to classify medical helpline calls as stroke or non-stroke. Calls were first transcribed using an automatic speech recognition model and then categorised by a text classification model trained as an ensemble of five individual models. We compared the performance of the model with that of call-takers using MH-1813 data from 2021.

@@

\begin{table}[t]
    \centering
    \caption[Performance metrics for stroke recognition.]{Performance metrics. Performance metrics for the model and call-takers on the test set. The model was trained on the training set, which included calls from 2015 to 2020. The call-takers were trained on the same data. The 2021 data were used for testing. The 2021 data without diagnostic categories were excluded}
    \label{tab_retrospective:table1-population-characteristics}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|ccccc}
        \toprule
                              & Training (112) & Training (MH-1813) & Validation & Test & 2021 w/o category \\

@@

The classification model outperformed the call-takers (\cref{tab_retrospective:table2-main-results}), with significant differences in all metrics (p < 0.0001). Excluding the 1-1-2 call line training data significantly degraded the model's performance (p < 0.0001), despite the domain mismatch with the MH-1813 call line test data. The performance on the 2021 calls without a diagnostic category was significantly worse than that of the test set regarding F1-score, sensitivity, false positive rate (FPR), and false omission rate (FOR) (p < 0.0001). The difference in positive predictive value (PPV) was not significant (p = 0.298).

The receiver operating characteristic (ROC) curve (\cref{fig_retrospective:figure1-roc-curve}, left) illustrates the potential to increase the sensitivity while maintaining a FPR lower than or equal to that of the call-takers. Similarly, the PPV sensitivity curve (\cref{fig_retrospective:figure1-roc-curve}, right) demonstrates that sensitivity can be improved while retaining a PPV higher than that of the call-takers. The framework can thus be tuned to a sensitivity of around 73\%, while still having a higher positive predictive value than the human call-taker (\cref{fig_retrospective:figure1-roc-curve}, right) The ensemble model outperformed the individual models regardless of the threshold, except for one that exhibited a slightly better sensitivity at a high FPR exceeding 1.5\%. The confusion matrices (\cref{fig_retrospective:figure2-prediction-confusion-matrices}) illustrate the performance differences in absolute numbers, with the model exhibiting more true positives and fewer false positives than the call-takers.

\begin{table}[t]
    \centering
    \caption[Overall stroke recognition performance of model compared to call-takers.]{Overall performance on MH-1813 test data, performance without 1-1-2 training data, and performance on data from 2021 without diagnostic categories as well as performance on MH-1813 test data based on demographic subgroups (age/sex) [mean (95\% CI)].}
    \label{tab_retrospective:table2-main-results}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|ccccc}
        \toprule
                    & F1-score [\%] $\uparrow$ & Sensitivity [\%] $\uparrow$ & PPV [\%] $\uparrow$ & \makecell{FOR [\%] $\downarrow$ \\ (1 - specificity)} & \makecell{FPR [\%] $\downarrow$ \\ (1 - NPV)} \\

@@

The model and call-takers exhibited significantly higher PPV and F1-score in men than in women (p < 0.0001) (\cref{tab_retrospective:table2-main-results}). The model significantly outperformed the call-takers on all metrics for each sex (p < 0.0001).
The model performed significantly better in the 65+ group than in the 18-64 year group regarding sensitivity, PPV, and F1-score (p < 0.0001). Similarly, the call-takers performed significantly better in the 65+ group than in the 18-64 group regarding PPV and F1-score (p < 0.0001). Finally, the model significantly outperformed the call-takers on all metrics in both age groups (p < 0.0001).

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{paper_retrospective/figure2.pdf}
    \caption[Prediction confusion matrices for stroke recognition.]{Prediction confusion matrices. Confusion matrices of predictions for call-takers and the model on the test set. Numbers for the model are given as the rounded mean over eleven runs.}
    \label{fig_retrospective:figure2-prediction-confusion-matrices}
\end{figure}

@@

We performed an occlusion analysis to evaluate the importance of individual words for both positive and negative classifier predictions \cref{tab_retrospective:table3-occlusion-analysis}
Among the words with a positive rank score, several words are synonymous with stroke, such as ``blood clot", ``haemorrhagic stroke", and ``stroke". Ambulances are rarely dispatched because the MH-1813 is not intended for emergencies. Therefore, a word like ``ambulance" may also be a strong indicator of call-taker recognition, which the model has learned to mimic. Additionally, most of the remaining words can be linked to stroke-related symptoms such as ``double vision", ``difficulties speaking", and ``hangs". Particularly, words describing the side of the body where symptoms occur ranked high (such as ``left", ``right", and ``side"). Finally, some words were also related to the sudden onset of symptoms (including ``suddenly" and ``minutes").

@@

\begin{table}[t]
    \centering
    \caption[Words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.]{English translation of words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.}
    \label{tab_retrospective:table3-occlusion-analysis}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|lr|lr}
        \toprule
                    & \multicolumn{2}{c|}{Positive ranking score} & \multicolumn{2}{c}{Negative ranking score} \\
        \midrule
                    & \multicolumn{2}{c|}{Stroke predictions, $D=1,897$} & \multicolumn{2}{c}{Non-stroke predictions, $D=342,133$} \\
        \midrule
                    & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ \\
        \midrule
        1.  & Ambulance & 1,680 & Tetanus & 4,378 \\        
        2.  & Blood clot & 895 & Pregnant & 8,749 \\        
        3.  & Left & 1,108 & Cut & 7,592 \\        
        4.  & Right & 1,050 & Bandage & 4,561 \\        
        5.  & Double vision & 84 & Amager (a location) & 23,776 \\        
        6.  & The words & 344 & O'clock & 94,436 \\        
        7.  & Suddenly & 783 & The emergency room & 42,809 \\        
        8.  & Arm & 709 & The police & 2,903 \\        
        9.  & Side & 1,139 & Swollen & 60,559 \\        
        10. & Stroke & 117 & Over the counter (OTC) & 4,641 \\        
        11. & Double & 113 & The neck & 30,151 \\        
        12. & Control & 134 & Fever & 112,586 \\        
        13. & Call & 39 & Prescription & 5,450 \\        
        14. & Numb & 94 & Centimetre & 12,026 \\        
        15. & Minutes & 763 & The knee & 8,875 \\        
        16. & Difficulties speaking & 44 & The pharmacy & 10,085 \\        
        17. & Haemorrhagic stroke & 133 & The stomach & 42,105 \\        
        18. & Hand & 297 & Psychiatric & 3,688 \\        
        19. & The ambulance & 521 & Pneumonia & 7,597 \\        
        20. & Slurred speech & 58 & Stomach pain & 10,551 \\        
        21. & Blood clots & 224 & Stool & 19,155 \\        
        22. & Fast & 663 & The ribs & 3,928 \\        
        23. & Express & 44 & Bleed & 10,501 \\        
        24. & Blood thinner & 259 & Bleeding & 24,313 \\        
        25. & Incoherent & 15 & Ribs & 2,941 \\        
        26. & Lopsided & 211 & Broken & 19,415 \\        
        27. & Reduced & 528 & Inflammation & 10,050 \\        
        28. & Hangs & 628 & Common cold & 8,127 \\        
        29. & Transient & 48 & Morning or morrow & 78,558 \\        
        30. & Not making sense & 14 & Swelling & 17,762 \\        
        \bottomrule
    \end{tabular}%
    }
\end{table}

@@

This study does not imply that a machine learning model can replace medical call-takers. The effectiveness of the model is fully reliant on the conversation between the call-taker and caller and the call-taker's ability to skillfully triage the patient. Instead, the model should be used as a supportive tool for call-takers in the decision-making process, contributing to a higher recognition of patients with stroke and potentially boosting the confidence of call- takers in their decisions. A similar machine learning model designed to predict cardiac arrest was tested in a randomised controlled trial (RCT) at CEMS.15 The results highlighted the necessity of incorporating input from call-takers. The machine learning model for cardiac arrest has subsequently been implemented in daily practice at CEMS, in a setup similar to the one presented in our study. However, implementation of our framework requires further investigation. The relative performance gap between call-takers and the model was larger in our study than in the cardiac arrest study,15 which may affect the results of a potential RCT.

To support future work and discussions beyond the scope of this study, \cref{app:supplementary-retrospective} includes the results of a simulation of a live implementation where call-takers are assumed to follow a set of fixed rules based on the output of the machine learning framework.
For instance, in one simulation call-takers are assumed to change any stroke negative to a positive, if the model predicts a positive. While the results of the simulation are encouraging, it is important to stress that it is not practically feasible to use a fixed rule set to overrule the call-taker. These results should only be seen as a preliminary indicator of a potential RCT. In practice, a nuanced set of guidelines should be developed over several iterations of implementation and testing.

@@

The model performed significantly better on men than on women. This could be attributed to several factors. First, the model may have learned to mimic call-takers with the same bias. Second, women may experience different and more challenging-to-identify symptoms than men \parencite{cite29, cite30}. Third, a higher prevalence of male patients with stroke was observed in the training data. Despite these potential sources of bias, the model exhibited less bias than call-takers did. That is, the relative performance improvements were higher for women than for men. This bias could be further reduced using advanced data augmentation and balanced data when training a machine learning model. However, such measures may degrade overall performance.

The improved sensitivity and PPV on the 65+ years group may be explained a higher prior probability of stroke for older patients and stronger evidence from the patient's medical history. The relatively high FOR and FPR for the 65+ group is likely to be a result of the much higher prevalence of stroke cases compared to the 18-64 year olds (0.85\% vs. 0.07\%). We did not have data to estimate potential bias related to race, ethnicity, language, accent, or dialects. Previous studies on speech recognition for call centres have indeed found that non-native speakers had a higher rate of transcription errors \parencite{han_deep_2017}. Since our model was trained on a representative - and therefore unbalanced - sample, we expect it to behave similarly. Future research should look to address these shortcomings, for example by utilising self"=supervised learning on massive amounts of diverse, unlabelled data covering multiple languages, accents, and dialects.

@@

The CEMS is responsible for providing prehospital telehealth services in the Capital Region of Denmark, with a catchment area of 1.9 million \parencite{cite17}. CEMS operates two call lines: the 1-1-2 emergency line, similar to 9-1-1 in the United States, intended for acute conditions. The other is the medical helpline 1813 (MH-1813, pronounced ``'eighteen-thirteen") intended for non- life-threatening conditions that cannot wait until a general practitioner is available.18

Call-takers for both lines, who are nurses, paramedics, or physicians, can dispatch ambulances. The condition suspected by the call-taker is categorised based on a predefined diagnostic index and stored in an electronic record using a computer-aided dispatch (CAD) system. The CAD records are associated with the Danish civil registration number (CPR number) \parencite{cite19} of the patient. The CPR number is a unique identification assigned to all Danish residents. It is used for interactions with health services and registries, enabling cross- referencing of the data sources used in this study. The call audio is recorded and stored separately from the CAD recordings using a telephone system.

@@

Stroke prevalence in calls made to the MH-1813 is lower than that in calls made to 1-1-2. Patients with stroke may exhibit different symptoms and symptom severity because MH- 1813 is meant for low-acuity incidents, leading to reduced recognition. In addition, MH-1813 call-takers dispatch high-priority transport less frequently, which may affect optimal treatment timing. Therefore, we focused on MH-1813 in this study.

@@

From the CAD medical records, we included all calls that could be matched to a corresponding audio file. for 1-1-2 and MH-1813 from 2015 to 2021 for patients older than 18. The CAD records were matched with the telephone call recordings based on the call start, call duration, and call-taker identity. Due to data incompleteness, and the way the audio data is stored, at CEMS, 2,730,199 contacts could not be matched to their corresponding audiofile, however, 2,361,178 contacts were successfully matched. We found no obvious pattern in the matched and unmatched calls and we included all calls with at matching audio file. Next, a call was regarded as a case of ground truth stroke positive when the CPR number in the CAD record matched that of a DanStroke record, and the patient-reported time of onset was close to the call start time. We allowed a window of 72 hours before and 24 hours after the call starts to account for uncertainty in recording stroke onset time. We excluded calls involving subarachnoid haemorrhage cases. Finally, we considered a call to be a call-taker stroke positive when the call-taker selected the stroke diagnostic category during the call and dispatched an ambulance with the appropriate level of response \parencite{cite20}. To ensure that the effect of the machine learning framework was not overestimated, we excluded calls where diagnostic category had not been registered from the test set. We still reported the population characteristics and model performance of this group of calls to assess potential bias introduced by excluding them. A data-flow diagram is included in \cref{app:supplementary-retrospective}. The resulting dataset is the largest dataset of audio files from stroke calls collected to date.

@@

As input for the classification model, each transcript was transformed into a fixed-size bag- of-words vector, which encoded the occurrence of word and character ($n$-grams) in a fixed vocabulary. The feature selection procedure is detailed in \cref{app:supplementary-retrospective}. The model was constructed as an ensemble \parencite{hansen_neural_1990} of five identical, independently trained models. Each consists of a stack of neural network layers commonly referred to as a multi-layer perceptron \parencite{rosenblatt_perceptron_1958}. The final layer has a single scalar output and applies a sigmoid nonlinearity to produce an output score between zero and one.

@@

As the threshold differed for each model in the ensemble, computing the ensemble output score as the average output score of the individual models would not be meaningful. Instead, we first subtracted the threshold from the output score in the logit space (before sigmoid nonlinearity) for each model to obtain the same threshold (0.5). Subsequently, we defined the ensemble output score as the average of the centred output scores. The exact equations are provided in \cref{app:supplementary-retrospective}.

@@

The speech recognition model was trained on 3,811 manually transcribed random calls (173 h) from the CEMS as part of a previous project.14 These calls exclusively originated from 1- 1-2 between 2015 and 2018, ensuring no overlap with the test data used for the text classification model. The model was trained using a connectionist temporal classification objective \parencite{graves_connectionist_2006}.

@@

We report the F1-score, sensitivity, PPV, false omission rate (FOR, equal to 1- negative predictive value), and false positive rate (FPR, equal to 1 - specificity). Due to the imbalanced nature of the dataset, the negative predictive value and specificity were greater than 99\% for all cases. We reported FOR and FPR instead because such large numerical values exhibit low relative variance, thereby obfuscating comparisons. Finally, we report the prediction confusion matrices, receiver operator characteristic (ROC) curve, and PPV- sensitivity curve, commonly known as the precision-recall curve. All results are reported with up to three significant digits.

We tested our results for statistical significance using approximate permutation tests. We used one-sided paired approximate permutation tests for model-to-model and model-to-call-taker comparisons when done on the same subset. For comparisons across different subsets (e.g. male vs. female) we used one-sided independent approximate permutation tests. We computed 95\% confidence intervals (CIs) using bootstrapping. \parencite{dwass_modified_1957,eden_validity_1933}. In our assessment, we accounted for random variation associated with model training by basing the means, tests, and CIs on the predictions of 11 randomly initialised training runs. Statistical significance was defined as a p-value of < 0.05.

@@

\subsection{Role of the funding source}

The grant providers had no role in the study design, data collection, analysis, interpretation, manuscript writing, or publication decision. Corti provided additional funds and technical expertise to develop the models. Corti was not financially compensated for this, and the project was part of its research initiatives, which were conducted in cooperation with several universities and the Innovation Fund Denmark. Corti owns the rights to the models and source code.

@@

The source code can be shared using a Creative Commons NC-ND 4.0, an international license, upon reasonable written request to the corresponding author, and requires a data use agreement.

>>>>

\section*{Abstract}
Advanced stroke treatment is time dependent and, therefore, relies on recognition by call-takers at prehospital telehealth services to ensure fast hospitalisation. This study aims to develop and assess the potential of machine learning in improving prehospital stroke recognition during medical helpline calls. We use calls from 1 January 2015 to 31 December 2020 in Copenhagen to develop a machine learning-based classification pipeline. Calls from 2021 are used for testing. Calls are first transcribed using an automatic speech recognition model and then categorized as stroke or non-stroke using a text classification model. Call-takers achieve a sensitivity of 52.7\% (95\% confidence interval 49.2-56.4\%) with a positive predictive value (PPV) of 17.1\% (15.5-18.6\%). The machine learning framework performs significantly better (p < 0.0001) with a sensitivity of 63.0\% (62.0-64.1\%) and a PPV of 24.9\% (24.3-25.5\%). Thus, a machine learning framework for recognising stroke in prehospital medical helpline calls may become a supportive tool for call-takers, aiding in early and accurate stroke recognition.

\section{Introduction}
Stroke is a leading cause of disability and death worldwide \parencite{cite1,cite2,cite3}. Effective treatment is time-sensitive, and an optimal outcome is more likely when treatment is administered within the first four and a half hours from stroke onset \parencite{cite4,cite5}. The gateway to ambulance transport and hospital admittance is through prehospital telehealth services, including emergency medical call centres, nurse advice call lines, and out-of-hours health services. In the pre-hospital setting, the use of mobile stroke units has made it possible to deliver advanced treatment faster \parencite{cite6,cite7}. As the mobile stroke unit is only dispatched to patients with a suspected stroke, the impact of mobile stroke unit is directly influenced by accurate call-taker recognition of stroke \parencite{cite6,cite7}. Call-takers who can rapidly and accurately recognise stroke are therefore crucial in facilitating prompt care in both pre-hospital and in-hospital settings.

@@

In this study, we use call recordings and registry data from the Copenhagen Emergency Medical Services (CEMS) and the Danish Stroke Registry (DanStroke) \parencite{cite16} from 2015 to 2020. We obtain call recordings from two call lines: the 1-1-2 emergency line and the medical helpline 1813 (MH-1813). We then fit a machine learning framework to classify medical helpline calls as stroke or non-stroke. Calls are first transcribed using an automatic speech recognition model and then categorised by a text classification model trained as an ensemble of five individual models. We compare the performance of the model with that of call-takers using MH-1813 data from 2021.

@@

\begin{table}[t]
    \centering
    \caption[Population characteristics for each data subset.]{Population characteristics for each data subset.}
    \label{tab_retrospective:table1-population-characteristics}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|ccccc}
        \toprule
                              & Training (112) & Training (MH-1813) & Validation & Test & 2021 w/o category \\

@@

The classification model outperformed the call-takers (\cref{tab_retrospective:table2-main-results}), with significant differences in all metrics (p < 0.0001, paired approximate permutation test). Excluding the 1-1-2 call line training data significantly degraded the model's performance (p < 0.0001, paired approximate permutation test), despite the domain mismatch with the MH-1813 call line test data. The performance on the 2021 calls without a diagnostic category was significantly worse than that of the test set regarding F1-score, sensitivity, false positive rate (FPR), and false omission rate (FOR) (p < 0.0001, independent approximate permutation test). The difference in positive predictive value (PPV) was not significant (p = 0.298, independent approximate permutation test).

The receiver operating characteristic (ROC) curve (\cref{fig_retrospective:figure1-roc-curve}, left) illustrates the potential to increase the sensitivity while maintaining a FPR lower than or equal to that of the call-takers. Similarly, the PPV-sensitivity curve (\cref{fig_retrospective:figure1-roc-curve}, right) demonstrates that sensitivity can be improved while retaining a PPV higher than that of the call-takers. The framework can thus be tuned to a sensitivity of around 73\%, while still having a higher positive predictive value than the human call-taker (\cref{fig_retrospective:figure1-roc-curve}, right). The ensemble model outperformed the individual models regardless of the threshold, except for one that exhibited a slightly better sensitivity at a high FPR exceeding 1.5\%. The confusion matrices (\cref{fig_retrospective:figure2-prediction-confusion-matrices}) illustrate the performance differences in absolute numbers, with the model exhibiting more true positives and fewer false positives than the call-takers.

\begin{table}[t]
    \centering
    \caption[Overall stroke recognition performance of model compared to call-takers.]{Overall performance on MH-1813 test data, performance without 1-1-2 training data, and performance on data from 2021 without diagnostic categories as well as performance on MH-1813 based on demographic subgroups (age/sex) [mean (95\% CI)]. NPV: negative predictive value, PPV: positive predictive value, FOR: false omission rate, CI: confidence interval.}
    \label{tab_retrospective:table2-main-results}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|ccccc}
        \toprule
                    & F1-score [\%] $\uparrow$ & Sensitivity [\%] $\uparrow$ & PPV [\%] $\uparrow$ & \makecell{FOR [\%] $\downarrow$ \\ (1 - specificity)} & \makecell{FPR [\%] $\downarrow$ \\ (1 - NPV)} \\

@@

The model and call-takers exhibited significantly higher PPV and F1-score in men than in women (p < 0.0001, independent approximate permutation test) (\cref{tab_retrospective:table2-main-results}). The model significantly outperformed the call-takers on all metrics for each sex (p < 0.0001, paired approximate permutation test).
The model performed significantly better in the 65+ group than in the 18-64 year group regarding sensitivity, PPV, and F1-score (p < 0.0001, independent approximate permutation test). Similarly, the call-takers performed significantly better in the 65+ group than in the 18-64 group regarding PPV and F1-score (p < 0.0001, independent approximate permutation test). Finally, the model significantly outperformed the call-takers on all metrics in both age groups (p < 0.0001, paired approximate permutation test).

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{paper_retrospective/figure2.pdf}
    \caption[Prediction confusion matrices for stroke recognition.]{Prediction confusion matrices. Confusion matrices of predictions for call takers and the model on the test set. Numbers for the model are given as the rounded mean over eleven runs.}
    \label{fig_retrospective:figure2-prediction-confusion-matrices}
\end{figure}

@@

We performed an occlusion analysis to evaluate the importance of individual words for both positive and negative classifier predictions \cref{tab_retrospective:table3-occlusion-analysis}. 
Among the words with a positive rank score, several words are synonymous with stroke, such as ``blood clot", ``haemorrhagic stroke", and ``stroke". Ambulances are rarely dispatched because the MH-1813 is not intended for emergencies. Therefore, a word like ``ambulance" may also be a strong indicator of call-taker recognition, which the model has learned to mimic. Additionally, most of the remaining words can be linked to stroke-related symptoms such as ``double vision", ``difficulties speaking", and ``hangs". Particularly, words describing the side of the body where symptoms occur ranked high (such as ``left", ``right", and ``side"). Finally, some words were also related to the sudden onset of symptoms (including ``suddenly" and ``minutes").

@@

\begin{table}[t]
    \centering
    \caption[Words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.]{English translation of words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively. For this analysis, we used the model with the median F1-score out of 11 randomly seeded runs.}
    \label{tab_retrospective:table3-occlusion-analysis}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|lr|lr}
        \toprule
                    & \multicolumn{2}{c|}{Positive ranking score} & \multicolumn{2}{c}{Negative ranking score} \\
        \midrule
                    & \multicolumn{2}{c|}{Stroke predictions, $D=1,897$} & \multicolumn{2}{c}{Non-stroke predictions, $D=342,133$} \\
        \midrule
                    & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ \\
        \midrule
        1.  & Ambulance & 1,680 & Tetanus & 4,378 \\        
        2.  & Blood clot & 895 & Pregnant & 8,749 \\        
        3.  & Left & 1,108 & Cut & 7,592 \\        
        4.  & Right & 1,050 & Bandage & 4,561 \\        
        5.  & Double vision & 84 & Amager (a location) & 23,776 \\        
        6.  & The words & 344 & O'clock & 94,436 \\        
        7.  & Suddenly & 783 & The emergency room & 42,809 \\        
        8.  & Arm & 709 & The police & 2,903 \\        
        9.  & Side & 1,139 & Swollen & 60,559 \\        
        10. & Stroke & 117 & Over the counter (OTC) & 4,641 \\        
        11. & Double & 113 & The neck & 30,151 \\        
        12. & Control & 134 & Fever & 112,586 \\        
        13. & Call & 39 & Prescription & 5,450 \\        
        14. & Numb & 94 & Centimetre & 12,026 \\        
        15. & Minutes & 763 & The knee & 8,875 \\        
        16. & Difficulties speaking & 44 & The pharmacy & 10,085 \\        
        17. & Haemorrhagic stroke & 133 & The stomach & 42,105 \\        
        18. & Hand & 297 & Psychiatric & 3,688 \\        
        19. & The ambulance & 521 & Pneumonia & 7,597 \\        
        20. & Slurred speech & 58 & Stomach pain & 10,551 \\        
        21. & Blood clots & 224 & Stool & 19,155 \\        
        22. & Fast & 663 & The ribs & 3,928 \\        
        23. & Express & 44 & Bleed & 10,501 \\        
        24. & Blood thinner & 259 & Bleeding & 24,313 \\        
        25. & Incoherent & 15 & Ribs & 2,941 \\        
        26. & Lopsided & 211 & Broken & 19,415 \\        
        27. & Reduced & 528 & Inflammation & 10,050 \\        
        28. & Hangs & 628 & Common cold & 8,127 \\        
        29. & Transient & 48 & Morning or morrow & 78,558 \\        
        30. & Not making sense & 14 & Swelling & 17,762 \\        
        \bottomrule
    \end{tabular}%
    }
\end{table}

@@

This study does not imply that a machine learning model can replace medical call-takers. The effectiveness of the model is fully reliant on the conversation between the call-taker and caller and the call-taker's ability to skillfully triage the patient. Instead, the model should be used as a supportive tool for call-takers in the decision-making process, contributing to a higher recognition of patients with stroke and potentially boosting the confidence of call-takers in their decisions. A similar machine learning model designed to predict cardiac arrest was tested in a randomised controlled trial (RCT) at CEMS \parencite{cite15}. The results highlighted the necessity of incorporating input from call-takers. The machine learning model for cardiac arrest has subsequently been implemented in daily practice at CEMS, in a setup similar to the one presented in our study. However, implementation of our framework requires further investigation. The relative performance gap between call-takers and the model was larger in our study than in the cardiac arrest study \parencite{cite15}, which may affect the results of a potential RCT.

To support future work and discussions beyond the scope of this study, the supplementary material includes the results of a simulation of a live implementation where call-takers are assumed to follow a set of fixed rules based on the output of the machine learning framework \cref{app:supplementary-retrospective}. For instance, in one simulation call-takers are assumed to change any stroke negative to a positive, if the model predicts a positive. While the results of the simulation are encouraging, it is important to stress that it is not practically feasible to use a fixed rule set to overrule the call-taker. These results should only be seen as a preliminary indicator of a potential RCT. In practice, a nuanced set of guidelines should be developed over several iterations of implementation and testing.

@@

The model performed significantly better on men than on women. This could be attributed to several factors. First, the model may have learned to mimic call-takers with the same bias. Second, women may experience different and more challenging-to-identify symptoms than men \parencite{cite29,cite30}. Third, a higher prevalence of male patients with stroke was observed in the training data. Despite these potential sources of bias, the model exhibited less bias than call-takers did. That is, the relative performance improvements were higher for women than for men. This bias could be further reduced using advanced data augmentation and balanced data when training a machine learning model. However, such measures may degrade overall performance.

The improved sensitivity and PPV on the 65+ years group may be explained a higher prior probability of stroke for older patients and stronger evidence from the patient's medical history. The relatively high FOR and FPR for the 65+ group is likely to be a result of the much higher prevalence of stroke cases compared to the 18-64 year olds (0.85\% vs. 0.07\%). We did not have data to estimate potential bias related to race, ethnicity, language, accent, or dialects. Previous studies on speech recognition for call centres have indeed found that non-native speakers had a higher rate of transcription errors \parencite{han_deep_2017}. Since our model was trained on a representative - and therefore unbalanced - sample, we expect it to behave similarly. Future research should look to address these shortcomings, for example by utilising self"-supervised learning on massive amounts of diverse, unlabelled data covering multiple languages, accents, and dialects.

@@

The CEMS is responsible for providing prehospital telehealth services in the Capital Region of Denmark, with a catchment area of 1.9 million \parencite{cite17}. CEMS operates two call lines: the 1-1-2 emergency line, similar to 9-1-1 in the United States, intended for acute conditions. The other is the medical helpline 1813 (MH-1813, pronounced ``'eighteen-thirteen") intended for non- life-threatening conditions that cannot wait until a general practitioner is available \parencite{cite18}.

Call-takers for both lines, who are nurses, paramedics, or physicians, can dispatch ambulances. The condition suspected by the call-taker is categorised based on a predefined diagnostic index and stored in an electronic record using a CAD system. The CAD records are associated with the Danish civil registration number (CPR number) \parencite{cite19} of the patient. The CPR number is a unique identification assigned to all Danish residents. It is used for interactions with health services and registries, enabling cross-referencing of the data sources used in this study. The call audio is recorded and stored separately from the CAD recordings using a telephone system.

@@

Stroke prevalence in calls made to the MH-1813 is lower than that in calls made to 1-1-2. Patients with stroke may exhibit different symptoms and symptom severity because MH-1813 is meant for low-acuity incidents, leading to reduced recognition. In addition, MH-1813 call-takers dispatch high-priority transport less frequently, which may affect optimal treatment timing. Therefore, we focused on MH-1813 in this study.

@@

From the CAD medical records, we included all calls that could be matched to a corresponding audio file for 1-1-2 and MH-1813 from 2015 to 2021 for patients older than 18. The CAD records were matched with the telephone call recordings based on the call start, call duration, and call-taker identity. Due to data incompleteness, and the way the audio data is stored, at CEMS, 2,730,199 contacts could not be matched to their corresponding audiofile, however, 2,361,178 contacts were successfully matched. We found no obvious pattern in the matched and unmatched calls and we included all calls with at matching audio file. Next, a call was regarded as a case of ground truth stroke positive when the CPR number in the CAD record matched that of a DanStroke record, and the patient-reported time of onset was close to the call start time. We allowed a window of 72 hours before and 24 hours after the call starts to account for uncertainty in recording stroke onset time. We excluded calls involving subarachnoid haemorrhage cases. Finally, we considered a call to be a call-taker stroke positive when the call-taker selected the stroke diagnostic category during the call and dispatched an ambulance with the appropriate level of response \parencite{cite20}. To ensure that the effect of the machine learning framework was not overestimated, we excluded calls where diagnostic category had not been registered from the test set. We still reported the population characteristics and model performance of this group of calls to assess potential bias introduced by excluding them. A data-flow diagram is included in \cref{app:supplementary-retrospective}. The resulting dataset is the largest dataset of audio files from stroke calls collected to date.

@@

As input for the classification model, each transcript was transformed into a fixed-size bag-of-words vector, which encoded the occurrence of word and character ($n$-grams) in a fixed vocabulary. The feature selection procedure is detailed in \cref{app:supplementary-retrospective}. The model was constructed as an ensemble \parencite{hansen_neural_1990} of five identical, independently trained models. Each consists of a stack of neural network layers commonly referred to as a multi-layer perceptron \parencite{rosenblatt_perceptron_1958}. The final layer has a single scalar output and applies a sigmoid nonlinearity to produce an output score between zero and one.

@@

As the threshold differed for each model in the ensemble, computing the ensemble output score as the average output score of the individual models would not be meaningful. Instead, we first subtracted the threshold from the output score in the logit space (before sigmoid nonlinearity) for each model to obtain the same threshold (0.5). Subsequently, we defined the ensemble output score as the average of the centred output scores. The exact equations are provided in \cref{app:supplementary-retrospective} in \cref{eq_retrospective:output_score, eq_retrospective:ensemble_prediction}.

@@

The speech recognition model was trained on 3,811 manually transcribed random calls (173 h) from the CEMS as part of a previous project \parencite{cite14}. These calls exclusively originated from 1- 1-2 between 2015 and 2018, ensuring no overlap with the test data used for the text classification model. The model was trained using a connectionist temporal classification objective \parencite{graves_connectionist_2006}.

@@

We report the F1-score, sensitivity, PPV, FOR (equal to 1- negative predictive value), and FPR (equal to 1 - specificity). Due to the imbalanced nature of the dataset, the negative predictive value and specificity were greater than 99\% for all cases. We reported FOR and FPR instead because such large numerical values exhibit low relative variance, thereby obfuscating comparisons. Finally, we report the prediction confusion matrices, ROC curve, and PPV-sensitivity curve, commonly known as the precision-recall curve. All results are reported with up to three significant digits.

We present the results with and without 1-1-2 training data, subgroup analyses based on age (18–64/65+) and sex (male/female), and call-takers performance. We also report the model performance on calls without a diagnostic category from the test year 2021 to assess potential data bias. We tested our results for statistical significance using approximate permutation tests. We used one-sided paired approximate permutation tests for model-to-model and model-to-call-taker comparisons when done on the same subset. For comparisons across different subsets (e.g. male vs. female) we used one-sided independent approximate permutation tests. We computed 95\% confidence intervals (CIs) using bootstrapping \parencite{dwass_modified_1957,eden_validity_1933}. In our assessment, we accounted for random variation associated with model training by basing the means, tests, and CIs on the predictions of 11 randomly initialised training runs. Statistical significance was defined as a p-value of less than 0.05.

@@

@@

The source code can be shared using a Creative Commons NC-ND 4.0, an international licence, upon reasonable written request to the corresponding author, and requires a data use agreement.
